The importance of measurement uncertainty in ecological management

### 2018-08-16 manuscripts/manuscript.Rmd
WIP

Imperfect information is ubiquitous in ecological management and conservation decision making. While the pressing concerns of global change have put the spotlight on *forecasting*, i.e. the uncertainty of the future [e.g. @Petchy2015], management decisions must also contend with uncertainty of the *present*: How many fish are in the sea today? What regions harbor the most fragile biodiversity? How have invasive species transformed native communities? A recognition of the importance of uncertainty has become widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierarchical Bayesian Modeling, along with convenient software tools for implementing these methods.  In sharp contrast to that shift, the treatment of uncertainty in decisions -- how the uncertainty in models should be reflected in scientific recommendations, remains far less developed, more qualitative, and under-served by available tools.  We illustrate how uncertainty in measurement has a dramatic negative impact on outcomes under a range of existing approaches, which can be avoided through the use of more formal approaches that have since been refined in fields such as robotics.  This insight both resolves a long-standing paradox in fisheries management and has important implications for how ecosystem management deals with uncertainty more generally.

Approaches for decision making under uncertainty in ecological systems can be divided into two camps: (1) those based upon "optimal control" solutions, and usually favored by natural resource economists, and (2) those based in heuristic methods such as scenario planning, resilience thinking, and precautionary rules of thumb, more commonly found in both ecological literature and actual practice [@Fischer2009; @Polasky2011].  While researchers have recognized the need to unify the transparent and quantitative algorithmic approach of optimal control with the greater complexity and uncertainty of real ecosystems that is acknowledged by heuristic methods, computational barriers to doing so have stymied this progress.  We illustrate how the limitations of these approaches have manifested in the example of fisheries management, and present a new approach that can combine the reality of measurement uncertainty and the rigor of optimization to resolve a long-standing paradox and suggest a more robust approach to management.  We compare the results of managing simulated fish stocks under (A) the existing optimal control approach, (B) a heuristic approach that is more typical of today's management practices, and (C) a newly proposed approach that provides a more rigorous treatment of uncertainty.

The challenge of making decisions under uncertainty is obviously not unique to fisheries, and the existing methods we critique here can be found in a wide range of population and ecosystem management challenges, including fire ecology [@Richards1999], invasive species [@Blackwood2010], and protecting biodiversity [@Dee2017].  Yet fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management more generally, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1954; @Gordon1954], thanks to its global relevance, long history, and readily available data.  Despite the generality of the models and issue of measurement uncertainty considered here, any application must be grounded in both the history and context of the particular decision problem. In light of this, we begin with a brief background on existing decision theory approaches common in fisheries management.

@@

For simplicity we will assume measurement uncertainty is normally distributed with standard deviation $\sigma_m$,  $\xi \sim \mathcal{N}(1, \sigma_m)$.  Given an estimated recruitment model along with this measurement of the stock size, the manager must choose the set of a harvest quotas $\lbrace h_t \rbrace$ for $t \in [0, 1, \infty]$ that maximizes expected long term (net) present value:

\begin{equation}
\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}
\end{equation}

A few subtleties arise in this seemingly simple problem statement that we must address.  First, given our discrete-time formulation of the recruitment process, it is necessary to decide if the measurement $y_t$ happens before or after harvest $h_t$: that is do we: measure, recruit, harvest, or measure, harvest, recruit?  Following convention in the optimal control literature (e.g. [@Reed1979; @Clark1990; @Sethi2005]), we will assume the latter; measurement occurs immediately before harvest, followed by recruitment.

A second subtly is in the nature of the decision problem itself: note that the manager in not myopic, thinking only about the next year, but rather considers the sequence of all future actions ${h_t}$.  While the manager will revisit this sequence each year in light of new observations, it is this ability of the decision problem to think ahead that allows it to tolerate short-term costs (e.g. reduced harvests in the current season) for larger future payoffs.  In this, ecological management is a game of chess, always thinking several moves ahead. This differs from the problem solved in theory as well as the practice of maximum sustainable yield, which seeks to identify only a fixed mortality $F_{MSY}$ applied for all time, rather than a *policy* of varying quotas $h_t$ depending on the stock assessment $y_t$.  Understanding this distinction is key both to understanding the differences between current theory and current best practice in fisheries management.

@@

It is important to remember that both existing methods we discuss and the POMDP approach proposed here can or are already applied to more complex models as well, including those with age or stage structure.  Likewise, while we consider the model and its parameters as "given," it should be understood to mean that this in practice comes as the result of some model choice and statistical estimation from historical data, frequently accounting for uncertainty in both the process and the measurements [e.g. @Dichmont2016].  As we shall see here, having included uncertainty in the model *estimation* in no way excuses us from having to also deal once again with that uncertainty in the decision process.

@@

Maximum Sustainable Yield (MSY) remains the basis of international law (including the UN,  IWC, IATTC, ICCAT, ICNAF; @Mace2001) and a familiar standard of management in terrestrial ecosystems as well as aquatic [@Clark1990].  Critics have for some time observed the limitations of harvesting at MSY in face of uncertainty in stock sizes and population dynamics [@Larkin1977; @Botsford1997].  Many US fisheries reflect this uncertainty through a series of adjustments that effectively reduce the target fishing mortality level to reflect this uncertainty.  Typically, a stock assessment model provides a best estimate of $B_{MSY}$ and corresponding mortality $F_{MSY}$ is used to define the stock Over-fishing Limit, (OFL). Based on this, a somewhat lower level is set as the Allowable Biological Catch (ABC), reflecting uncertainty in the stock assessment. To reflect possible uncertainty between reported and actual catch, the ABC is reduced somewhat further to define the Total Allowable Catch, (TAC), which forms the basic unit of management for many such fisheries.  To reflect this process, our analysis will also consider policies in which the harvest quota is set at 80% of the level expected under the MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$. To distinguish this approach from MSY, we will refer to this approach as a TAC policy.  This more closely represents a heuristic [e.g. @Hilborn2010] or resilience-based approach [sensu @Fischer2009] than an optimization-based policy.  Importantly, this approach shares the fundamentally stationary assumptions of an MSY policy by defining a constant mortality rather than a dynamic policy.  Thus, despite being more cautious overall and generating lower economic yield than a constant escapement policy, TAC policies continue to harvest at non-zero rates even if a stock falls below $B_{MSY}$, while the constant escapement policy does not.

<!--- FIXME MSE discussion HERE!! -->

@@

So far, attempts to provide a more formal basis for managing uncertainty than the heuristic adjustment of catch limits described above have largely foundered in a series of paradoxes.  Without uncertainty, the theoretical policies are quite intuitive: The result derived by @Clark1973 for an optimal dynamic strategy to replace the equilibrium solution of MSY can be summarized as: If the stock is most productive at $B_{MSY}$, then obtain $B_{MSY}$ as quickly as possible. Thus, at any level below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, harvests greater than $H_{MSY}$ will be needed to bring the stock back to $B_{MSY}$.  (This intuition must be adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).  This strategy is known as *constant escapement* (CE), since a constant sock size $B_{MSY}$ escapes harvest each year.  Because CE and MSY converge to the *same long-term biomass and same long-term average yield*, CE can appear to be merely a more formal justification for MSY.  However, in practice the two approaches are very different: becuase natural populations fluctuate year-to-year, under CE, a different harvest quota must be set each year to achieve the constant escapement.

In the face of a naturally fluctuating population, previous arguments made by @Gordon1954; @Schaefer1954 or @Clark1973 no longer hold.  The optimal harvest of a population under a model such as the stochastic Gordon Schaefer, Eq \eqref{Gordon-Schaefer}, is an example of a Markov Decision Problem (MDP). This is not to be confused with the the problem of *estimating* parameters or posteriors of a Markov process model, e.g. through typical Bayesian analysis -- recall this is a decision problem, in which the model has already been estimated and now we are tasked with choosing a sequence of future harvests based on that model.  Such MDP problems are commonly referred to in ecological literature by its solution method, Stochastic Dynamic Programming (SDP) [@Marescot2013], widely used in both ecological management and animal behavior [@Mangel1985; @Mangel1988].  However, thanks to a mathematical proof provided by @Reed1979, fisheries managers have largely been able to bi-pass the need for SDP.  But this convenient result also opened the door to paradox of uncertainty that would persist for the next four decades.

@@

Despite these notes of caution, both @Clark1986 and @Sethi2005 ultimately attempt to rationalize this counter-intuitive conclusion rather than reject it.  Intuitively, uncertainty in the population growth model (stochasticity) or measurement error, would seem to justify harvesting fewer fish. Instead, existing optimal control theory has presented only the paradoxical results that uncertainty either shoud not matter at all, or that we should increase rather than decrease harvests in response to increased uncertainty. Numerous other attempts have been made to resolve this paradox through alternative assumptions,tThe most common being to assume from the outset to look only for 'constant-escapement' type polcies [@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003]. Owing to the "mathematical difficulty" Clark first observed, none have attempted a direct solution which we will present here.

In practice, TAC-managed fisheries are more common that CE-managed fisheries, at least in the United States. CE-based regulation is limited primarily to salmon, species for which population estimates may be more precise and thus closer to Reed's assumption.  We are not aware of any fishery that has put Clark's Paradox into practice, harvesting more in response to increased uncertainty.  As we shall illustrate, this is particularly fortunate; though approaches that properly account for both forms of uncertainty can perform even better.

@@

Why does the introduction of measurement uncertainty make the decision problem so much harder?  Consider the thought experiment of managing a fish stock where in each of the previous four years stock assessments have put the population size at, say, 51, 54, 49, 51, (say, measured in percent of carrying capacity, $K$). Under a constant escapement target of 50, the harvest quota would be 1, 4, 0, 1, respectively.  Note that this control rule depends only on the measurement for the year in question, which is assumed to be measured without error. If we then measure a stock size of 75, an assumption that this measurement is made without error would justify the sudden increase to a quota of 25. Yet if we admit the possibility of measurement error, this sudden bump starts to look very suspicious in light of all the previous observations.  Perfect measurements let us exploit the Markov property of the model -- the most recent measurement tells us all we need to know.  In contrast, the introduction of measurement uncertainty breaks this Markovian assumption: a rational decision maker would suspect this 75 to be an over-estimate based on prior observations. Under imperfect observations, we no longer have a Markov Decision Process (MDP) over the observed state space.  Rather, we have a Partially Observed Markov Decision Process (POMDP). A POMDP policy cannot be expressed as a policy function at all (much less as policy function with a single constant-escapement target, as much previous work has assumed).  Instead, a POMDP policy can be thought of as depending not only on the most recent observation, but the recent observation combined with a *prior belief* about what state the system is in, which itself is determined by all past observations. In this manner, the POMDP can consider how much weight to give the most recent observation based on how it compares to prior beliefs.

Algorithmic innovations from the robotics community have now made it possible to numerically solve non-trivial POMDP problems. The need to consider all past observations becomes particularly demanding given the forward-looking nature of a decision problem.  Not only must we consider all possible sequences of future actions over all possible sequences of future states, but we must add to that all possible sequences of future measurements. The computational complexity involved has historically limited POMDP applications to contexts with only a handful of possible states and actions [e.g. @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b], insufficient to capture fisheries models appropriately.  Meanwhile, methods to solve decision problems with imperfect information have advanced steadily in the field of robotics and artificial intellengence.  Almost contemporaneous with Reed's work,  papers by @Smallwood1973 and @Sondik1978 laid mathematical foundations for algorithms that could efficiently solve POMDP problems [e.g. @Kaelbling1998; @Pineau2003].  More recently, driven by demands in areas such as autonomous vehicle decision making, newer point-based approximation algorithms such as SARSOP [@Kurniawati2008] have made it possible to solve POMDPs with 100s of states and actions.  We have adapt this algorithm for application to the fisheries context, and provide an efficient implementation as a preliminary R package [@sarsop-pkg] which can be used to replicate and further explore results presented here. Details of the analysis presented here including annotated code to reproduce and further explore all of the following results is provided in the supplementary material.

@@

We consider the average fish biomass across 100 replicate simulations under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.  Each suite of simulations is considered under three sequentially higher measurement error regimes, from $\sigma_m = 0$, $\sigma_m = 0.1$, and $\sigma_m = 0.15$, as indicated, while stochastic recruitment (environmental noise) is set to a moderate $\sigma_g = 0.15$.  In the absence of either environmental noise or measurement error POMDP, CE, and MSY would converge to stock at the $B_{MSY}$, while TAC would maintain the stock at a slightly higher level (see Supplemental material, Figure S1).  The introduction of stochastic growth has a significant negative impact on the TAC strategy, whith a mean biomass significantly lower than $B_{MSY}$ (first panel, Figure \ref{sims}; this impact is even more severe for MSY, see Figure S2).  Without measurement error, the CE and POMDP strategies are nearly identical with both approximately maintaining the stock at $B_{MSY}$ despite the significant environmental stochasticity.  As measurement error increases in the subsequent panels, CE and TAC strategies perform increasingly poorly, while the POMDP continues to maintain the average stock close to $B_{MSY}$.  Notably, CE is significantly more impacted by measurement uncertainty than TAC, with CE averaging even lower biomass than TAC under moderate measurement uncertainty.  These results confirm that the precautionary approach represented by the TAC does indeed prove more robust to the problem of measurement error than the optimization solution represented by CE, but is less robust to environmental error. In contrast, the POMDP approach successfully handles both sources of uncertainty, maintaining higher stocks, near $B_{MSY}$ level, and producing the highest yield.

@@

Using modern algorithms we have been able to both crack the nut of measurement uncertainty in the optimal management of marine fisheries and debunk a long-standing paradox: that increased measurement uncertainty should be met with larger harvests.  We have demonstrated that both existing optimal control approaches and existing precautionary approaches can result in over-exploitation and long term declines of fish stocks, as well as reduced economic value, while in contrast, the POMDP approach is able to deliver both ecological and economic recovery despite uncertainty in measurements.  These results have important implications for both global fisheries and ecological management more generally.  In marine fisheries, evidence of such over-exploitation has already been well documented [@Worm2006; @Worm2009; @Costello2016]. Our results suggest that measurement uncertainty may contribute to this pattern, as such errors lead to over-exploitation under current best practices such as MSY, TAC or CE, which have previously been thought to be sufficient to ensure the recovery of stocks world-wide [@Costello2016].

Our results also have important consequences for ecological management more generally. The study and application of management strategies such as MSY and CE underpin much of conservation and natural resource management in forest management and many other terrestrial ecosystems [e.g. @Clark2000]. Decision making under uncertainty is a central challenge for ecology and conservation biology [@Polasky2011], which we have long attempted to address through either simplified optimal control models or more heuristic approaches such as the precautionary principle and resilience thinking [@Fischer2009]. We have illustrated how both of these approaches can fail to provide an adequate strategy when faced with measurement uncertainty. To avoid this pitfall, a strategy must be dynamic to make best use of the available information without assuming that information is perfect.  Classical optimal control approaches (such as CE) can fail due to misplaced confidence in measurements even when they are significantly at odds with prior expectations. Similarly, the TAC approach also fails due to its reliance on a simple one-size-fits-all strategy, not reducing catch quotas sufficiently when stocks are too low but also failing to increase them when stocks are higher.   It is precisely this ability to adapt the policy in response to available information, as illustrated in Figure \ref{policy}, that makes the POMDP approach successful. Nevertheless, it may be tempting to see these results as either trivial or impractical, but a closer inspection reveals they are neither.

>>>>

Imperfect information is ubiquitous in ecological management and conservation decision making. While the pressing concerns of global change have put the spotlight on *forecasting*, i.e. the uncertainty of the future [e.g. @Petchy2015], management decisions must also contend with uncertainty of the *present*: How many fish are in the sea today? What regions harbor the most fragile biodiversity? How have invasive species transformed native communities? A recognition of the importance of uncertainty has become widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierarchical Bayesian Modeling, along with convenient software tools for implementing these methods to build better models.  As we estimate ever more sophisticated models, it may be easy to forget that a model is not itself a decision, particularly when the estimated model includes considerable uncertainty about future outcomes.  Knowing the rules of the chess board is not the same thing as knowing how to win. A model only gives us the rules of the game. The science of decision theory helps us know how to play.  Yet while our models have become much more sophisticated, decision theory in ecological management frequently ignores the very uncertainty we so carefully include when estimating the model.  Previous shortcuts in decision theory have led to paradoxical results, such as the suggestion that increased uncertainty in fisheries stock sizes should justify unchanged or even increased harvest rates.  Borrowing decision theory methodology first developed in fields such as robotics, we are able to resolve this paradox, underscoring the importance of appropriately accounting for uncertainty in decisions just as we already do in models. This insight has important implications in both fisheries and how ecosystem management deals with uncertainty more generally.

Approaches for decision making under uncertainty in ecological systems can be divided into two camps: (1) those based upon "optimal control" solutions, and usually favored by natural resource economists, and (2) those based in heuristic methods such as scenario planning, resilience thinking, and precautionary rules of thumb, more commonly found in both ecological literature and actual practice [@Fischer2009; @Polasky2011].  While researchers have recognized the need to unify the transparent and quantitative algorithmic approach of optimal control with the greater complexity and uncertainty of real ecosystems that is acknowledged by heuristic methods, computational barriers to doing so have stymied this progress.  We illustrate how the limitations of these approaches have manifested in the example of fisheries management, and present a new approach that can combine the reality of measurement uncertainty and the rigor of optimization to resolve a long-standing paradox and suggest a more robust approach to management.  We compare the results of managing simulated fish stocks under (A) the existing optimal control approach, (B) a heuristic approach that is more typical of today's management practices, and (C) a unified approach that provides a more rigorous treatment of uncertainty.

The challenge of making decisions under uncertainty is obviously not unique to fisheries, and the existing methods we critique here can be found in a wide range of population and ecosystem management challenges, including fire ecology [@Richards1999], invasive species [@Blackwood2010; @Shea1998], disease outbreaks [@Shea2014; @Li2017], and protecting biodiversity [@Dee2017; @Iacona2017].  Yet fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management more generally, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1954; @Gordon1954], thanks to its global relevance, long history, and readily available data.  Despite the generality of the models and issue of measurement uncertainty considered here, any application must be grounded in both the history and context of the particular decision problem. In light of this, we begin with a brief background on existing decision theory approaches common in fisheries management.

@@

For simplicity we will assume measurement uncertainty is normally distributed with standard deviation $\sigma_m$,  $\xi \sim \mathcal{N}(1, \sigma_m)$.  Given an estimated population model along with this (uncertain) measurement of the stock size, the manager must choose the set of a harvest quotas $\lbrace h_t \rbrace$ for $t \in [0, 1, \infty]$ that maximizes expected long term utility of the stock:

\begin{equation}
\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot U(x_t, h_t) \right \}
\end{equation}

For convenience of presentation but without loss of generality, we will assume this utility is simply equal to the harvest itself; in economic terms, a fixed price for fish without cost on effort:

\begin{equation}
 U(x_t, h_t) = \min(h_t, x_t)
\end{equation}

where the $\min$ function ensures no additional reward for attempting to harvest more than the entire stock.  Existing optimal control techniques (@Reed1979) prove these results are not dependent on the particular details of this utility.  Using this form allows a more direct comparison to approaches such as maximum sustainable yield (MSY), that do not include an economic model but merely maximize long-term catch. In principle, utility could also reflect ecosystem services or other intrinsic values for fish left in the sea and not merely the market for consumption [@Halpern2013].

The solution to this decision problem is conditional on a population model for the underlying ecological process, describing probability the system moves to state $x_{t+1}$ given that it was in state $x_t$ and the manager selected a harvest $h_t$, which we discuss below.  Such a model would typically be estimated and parameterized from available data, often (but not necessarily) using a hierarchical Bayesian approach that acknowledges the existence of uncertainty in both measurements and process [e.g. see @Dichmont2016 for a discussion of how uncertainties are frequently included in stock assessment models].  However, accounting for measurement uncertainty when going from data to model does not mean we have also accounted for this uncertainty when going from model to decision.  For example, a harvest decision policy based upon maximum sustainable yield simply depends on the stock size which maximizes the expected growth rate of the model.  Measurement uncertainty in model estimation can indeed influence the position of that maximum, but as we shall see, that strategy is only optimal in the absence of any uncertainty.

There is one further subtlety arising in this seemingly simple problem statement that we must address. Given our discrete-time formulation of the recruitment process, it is necessary to decide if the measurement $y_t$ happens before or after harvest $h_t$: that is do we: measure, recruit, harvest, or measure, harvest, recruit?  Following convention in the optimal control literature (e.g. [@Reed1979; @Clark1990; @Sethi2005]), we will assume the latter; measurement occurs immediately before harvest, followed by recruitment. This assumption has occasionally been reversed to introduce some additional uncertainty from stochastic recruitment immediately prior to setting the harvest quota [e.g. @Clark1986; @Weitzmann2001], which appears to mimic uncertainty in measurements but does not allow uncertainty to accumulate since the measurement step still occurs without error.

The optimal strategy must think not only about the next year, but rather considers the sequence of all future actions ${h_t}$, under all possible realizations of the underlying model.  This strategy must also acknowledge that the manager will receive new information each year, and may need to modify actions accordingly.  It is this ability of the decision problem to think ahead that allows it to tolerate short-term costs (e.g. reduced harvests in the current season) for larger future payoffs.  When Eq (1) is ignored, such optimal control problems are known as Markov Decision Processes or MDPs.  These are more commonly referred to in the ecological literature by their solution method of Stochastic Dynamic Programming [SDP; @Marescot2013], and widely used in both ecological management and animal behavior [@Mangel1985; @Mangel1988]. The addition of measurement uncertainty in Eq (1) turns this problem into a Partially Observed Markov Decision Process (POMDP), for which solution methods have not been widely available for large problems until more recently.  We discuss these decision methods in more detail below.

@@

It is important to remember that all three decision methods we compare here will use this same model as given.  In the case of numerical simulations, we will simply pick reasonable fixed parameter values for the model above.  To illustrate the application to empirical data, we will first estimate these parameters using a hierarchical Markov Chain Monte Carlo (MCMC), and then use the resulting estimates to drive all three decision models.  As such, the details of how the model is estimated are immaterial to the differences in performance between these decision methods.  In all cases, the model is considered fixed.  Technically, it would be possible to re-estimate the model parameters given each new observation during the decision process.  Adaptations of the decision processes we consider here include this possibility for learning about the model during the decisions. Some include the possibility of taking knowingly bad actions, like suddenly increasing harvest rates, to improve learning over parameters.  We investigate these in subsequent work [@Memarzadeh2018].  For better comparison with established theory, here, the model is considered fixed. In the case of the simulations, the model is known without error (though still stochastic in nature), such that additional observations cannot improve the parameter estimates in any event.  Despite that obvious simplification, we shall illustrate how existing decision methods fail to account for measurement uncertainty appropriately.  Even with these simple models, decisions are complex, and knowing the rules of the game is not the same as knowing how to win.

@@

Maximum Sustainable Yield (MSY) remains the basis of international law (including the UN,  IWC, IATTC, ICCAT, ICNAF; @Mace2001) and a familiar standard of management in terrestrial ecosystems as well as aquatic [@Clark1990].  Critics have for some time observed the limitations of harvesting at MSY in face of uncertainty in stock sizes and population dynamics [@Larkin1977; @Botsford1997].  Many US fisheries reflect this uncertainty through a series of adjustments that effectively reduce the target fishing mortality level to reflect this uncertainty.  Typically, a stock assessment model provides a best estimate of $B_{MSY}$ and corresponding mortality $F_{MSY}$ is used to define the stock Over-Fishing Limit, (OFL). Based on this, a somewhat lower level is set as the Allowable Biological Catch (ABC), reflecting uncertainty in the stock assessment. To reflect possible uncertainty between reported and actual catch, the ABC is reduced somewhat further to define the Total Allowable Catch, (TAC), which forms the basic unit of management for many such fisheries.  To reflect this process, our analysis will also consider policies in which the harvest quota is set at 80% of the level expected under the MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$. To distinguish this approach from MSY, we will refer to this approach as a TAC policy.  This more closely represents a heuristic [e.g. @Hilborn2010] or resilience-based approach [sensu @Fischer2009] than an optimization-based policy.  Importantly, this approach shares the fundamentally stationary assumptions of an MSY policy by defining a constant mortality rather than a dynamic policy.  Thus, despite being more cautious overall and generating lower economic yield than a constant escapement policy, TAC policies continue to harvest at non-zero rates even if a stock falls below $B_{MSY}$, while the constant escapement policy does not.

It is important to note that actual management strategies are quite diverse.  Our focus on comparisons TAC, MSY, and Constant Escapement (CE, also refereed to as MEY, see below) reflects only the subset of current practices.  Another increasingly common approach to address uncertainty in both process and models is known as Management Strategy Evaluation [MSE, @Smith1994;  @Dorner2009; @Bunnefeld2011; @Punt2016].  MSE performs forward simulations under a suite of pre-selected candidate strategies to determine which strategy achieves the best expected outcomes. Crucially, by limiting the analysis to a suite of candidate strategies over a finite number of replicate simulations, this approach is able to accommodate substantially greater complexity and uncertainty than has been possible with more formal optimal control approaches we discuss below.  Optimal control solutions rely on dynamic programming to consider all possible strategies (all possible sequences of harvest under all possible realizations of stock dynamics).  As such, MSE can be viewed as an approximation to the optimal control solutions considered below, a notion that has been formalized in the literature on approximate dynamic programming [@Nicol2011].

@@

So far, attempts to provide a more formal basis for managing uncertainty than the heuristic adjustment of catch limits described above have largely foundered in a series of paradoxes.  Without uncertainty, the theoretical policies are quite intuitive: The result derived by @Clark1973 for an optimal dynamic strategy to replace the equilibrium solution of MSY can be summarized as: If the stock is most productive at $B_{MSY}$, then obtain $B_{MSY}$ as quickly as possible. Thus, at any level below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, harvests greater than $H_{MSY}$ will be needed to bring the stock back to $B_{MSY}$.  (This intuition must be adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).  This strategy is known as *constant escapement* (CE), since a constant sock size $B_{MSY}$ escapes harvest each year.  Because CE and MSY converge to the *same long-term biomass and same long-term average yield*, CE can appear to be merely a more formal justification for MSY.  However, in practice the two approaches are very different: because natural populations fluctuate year-to-year, under CE, a different harvest quota must be set each year to achieve the constant escapement.  Because CE is derived as the economic optimum (assuming perfect measurements), it is sometimes referred to as Maximum Economic Yield, [MEY; @Burgess2018], or as Rights Based Fisheries Management [RBFM, @Costello2016].

In the face of a naturally fluctuating population, previous arguments made by @Gordon1954; @Schaefer1954 or @Clark1973 no longer hold.  As noted above, the optimal harvest of a population under a model such as the stochastic Gordon Schaefer, Eq \eqref{Gordon-Schaefer}, is an example of a Markov Decision Problem (MDP) which must be solved with stochastic dynamic programming. Thanks to a mathematical proof provided by @Reed1979, fisheries managers have largely been able to bi-pass such computational effort.  However, this convenient result also opened the door to paradox of uncertainty that would haunt ecological management for at least the next four decades.

@@

Despite these notes of caution, both @Clark1986 and @Sethi2005 ultimately attempt to rationalize this counter-intuitive conclusion rather than reject it.  Intuitively, uncertainty in the population growth model (stochasticity) or measurement error, would seem to justify harvesting fewer fish. Instead, existing optimal control theory has presented only the paradoxical results that uncertainty either should not matter at all, or that we should increase rather than decrease harvests in response to increased uncertainty. Numerous other attempts have been made to resolve this paradox through alternative assumptions, the most common being to assume from the outset to look only for 'constant-escapement' type policies [@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003]. Owing to the "mathematical difficulty" Clark first observed, none have attempted a direct solution which we will present here.

In practice, TAC-managed fisheries are more common that CE-managed fisheries, at least in the United States. CE-based regulation is limited primarily to salmon, species for which population estimates may be more precise and thus closer to Reed's assumption.  We are not aware of any fishery that has put Clark's Paradox into practice, harvesting more in response to increased uncertainty.  As we shall illustrate, this is fortunate; though approaches that properly account for both forms of uncertainty can perform even better.

@@

Why does the introduction of measurement uncertainty make the decision problem so much harder?  Consider the thought experiment of managing a fish stock where in each of the previous four years stock assessments have put the population size at, say, 51, 54, 49, 51, (say, measured in percent of carrying capacity, $K$). Under a constant escapement target of 50, the harvest quota would be 1, 4, 0, 1, respectively.  Note that this control rule depends only on the measurement for the year in question, which is assumed to be measured without error. If we then measure a stock size of 75, an assumption that this measurement is made without error would justify the sudden increase to a quota of 25. Yet if we admit the possibility of measurement error, this sudden bump starts to look very suspicious in light of all the previous observations.  Perfect measurements let us exploit the Markov property of the model -- the most recent measurement tells us all we need to know.  In contrast, the introduction of measurement uncertainty breaks this Markovian assumption: a rational decision maker would suspect this 75 to be an over-estimate based on prior observations. Under imperfect observations introduced by Eq (1), we no longer have a Markov Decision Process (MDP) over the observed state space.  Rather, we have a Partially Observed Markov Decision Process (POMDP). A POMDP policy cannot be expressed as a policy function at all (much less as policy function with a single constant-escapement target, as much previous work has assumed).  Instead, a POMDP policy can be thought of as depending not only on the most recent observation, but the recent observation combined with a *prior belief* about what state the system is in, which itself is determined by all past observations. In this manner, the POMDP can consider how much weight to give the most recent observation based on how it compares to prior beliefs.

Algorithmic innovations from the robotics community have now made it possible to numerically solve non-trivial POMDP problems. The need to consider all past observations becomes particularly demanding given the forward-looking nature of a decision problem.  Not only must we consider all possible sequences of future actions over all possible sequences of future states, but we must add to that all possible sequences of future measurements. The computational complexity involved has historically limited POMDP applications to contexts with only a handful of possible states and actions [e.g. @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b], insufficient to capture fisheries models appropriately.  Meanwhile, methods to solve decision problems with imperfect information have advanced steadily in the field of robotics and artificial intelligence.  Almost contemporaneous with Reed's work,  papers by @Smallwood1973 and @Sondik1978 laid mathematical foundations for algorithms that could efficiently solve POMDP problems [e.g. @Kaelbling1998; @Pineau2003].  More recently, driven by demands in areas such as autonomous vehicle decision making, newer point-based approximation algorithms such as SARSOP [@Kurniawati2008] have made it possible to solve POMDPs with 100s of states and actions.  We have adapt this algorithm for application to the fisheries context, and provide an efficient implementation as a preliminary R package [@sarsop-pkg] which can be used to replicate and further explore results presented here. Details of the analysis presented here including annotated code to reproduce and further explore all of the following results is provided in the supplementary material.

@@

We consider the average fish biomass across 100 replicate simulations under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.  Each suite of simulations is considered under three sequentially higher measurement error regimes, from $\sigma_m = 0$, $\sigma_m = 0.1$, and $\sigma_m = 0.15$, as indicated, while stochastic recruitment (environmental noise) is set to a moderate $\sigma_g = 0.15$.  In the absence of either environmental noise or measurement error POMDP, CE, and MSY would converge to stock at the $B_{MSY}$, while TAC would maintain the stock at a slightly higher level (see Supplemental material, Figure S1).  The introduction of stochastic growth has a significant negative impact on the TAC strategy, with a mean biomass significantly lower than $B_{MSY}$ (first panel, Figure \ref{sims}; this impact is even more severe for MSY, see Figure S2).  Without measurement error, the CE and POMDP strategies are nearly identical with both approximately maintaining the stock at $B_{MSY}$ despite the significant environmental stochasticity.  As measurement error increases in the subsequent panels, CE and TAC strategies perform increasingly poorly, while the POMDP continues to maintain the average stock close to $B_{MSY}$.  Notably, CE is significantly more impacted by measurement uncertainty than TAC, with CE averaging even lower biomass than TAC under moderate measurement uncertainty.  These results confirm that the precautionary approach represented by the TAC does indeed prove more robust to the problem of measurement error than the optimization solution represented by CE, but is less robust to environmental error. In contrast, the POMDP approach successfully handles both sources of uncertainty, maintaining higher stocks, near $B_{MSY}$ level, and producing the highest yield.

@@

Using modern algorithms we have been able to both crack the nut of measurement uncertainty in the optimal management of ecological populations such as marine fisheries, debunking a long-standing paradox created by previous literature which has argued that increased measurement uncertainty should be met with equal or larger harvests [e.g. @Reed1979; @Clark1986; @Sethi2005].  By using powerful algorithms developed first the robotics literature to solve POMDPs [@Kurniawati2008], we have been able to tackle this optimal control problem directly rather than resorting to the simplifying assumptions used in previous work.  Our solution confirms the intuition of the precautionary principle in a quantitatively precise form: under most (but not all) conditions, *greater uncertainty results in smaller optimal harvests*.  Despite frequently proposing smaller harvests than existing approaches, the POMDP solution achieves higher long-term economic value as well as higher average stock sizes over the long term.

This lesson has clear relevance in the context of fisheries management.  Here, practice is already wiser than theory, where the optimal control based on the CE results of @Clark1973 and @Reed1979 is limited to fisheries such as salmon where estimates in stock size may be less significant than in purely marine species. To our knowledge, no fishery has intentionally put the more aggressive harvest solutions of @Clark1986 and @Sethi2005 into practice.  Yet this does not lift the specter of these methods, where CE-based predictions feature prominently in global fisheries analyses as the economically optimal strategy and the fastest route to recovery [@Costello2016; @Burgess2018]. Our results raise serious questions about the optimistic forecasts predicted under these CE-based policies. Meanwhile, MSY remains the dominate standard international law [@Mace2001], and many well-managed US fisheries rely on a version of the TAC based precautionary adjustment to MSY to comply with catch limits under Magnuson-Stevens Act.  Our results show that these strategies are less sensitive to measurement error, but not immune.  Comparisons against more heuristic decision methods are more difficult: management strategy evaluation [MSE; @Smith1994; @Punt2016] will depend on precisely what management scenarios are considered.  Under sufficient uncertainty, both the existing optimal and precautionary approaches can lead to declines in biomass and economic returns which could be avoided by POMDP-based management.
